A Temporal Logic Based Framework for Intrusion Detection

### 2023-09-04 Paper/main.tex
update

As we can see, rules are parameterized by an EAGLE formula (of type $\underline{\text{Form}}$), which means that we will be able to write EAGLE formulas such as $\text{Always}(\text{Eventually}(x>0))$. Also, the $\text{Always}$ operator is defined as maximal solution of the equation $X=F\land \bigcirc X$ and the $\text{Sometime}$ operator represents the minimal solution to the equation $X=F\lor \bigcirc X$.

@@

Given a trace $\sigma$ and a specification $D \, O$, satisfaction is defined as follows:
\[
\sigma \models D \, O \text{ iff } \forall (\underline{\text{mon}}\ N = F) \in O.\ \sigma ,1 \models_D \, F
\]
In other words, a trace satisfies a specification if it fulfills each monitored formula when monitoring from position 1, which is the initial state. An \textbf{important} note is that, given a trace $\sigma=\alpha_{1},\ldots ,\alpha_{n}$ the index $i$ of the trace can become $0$ (before the first state) and $n+1$, thus going beyond the limits; both of these situations result in rule applications evaluating to either true if maximal or false if minimal, without first taking the rules body into account. The definition of the satisfaction relation $\models_D \subseteq (Trace \times \mathbb{N}) \times \underline{\text{Form}}$, for a set of rule definitions $D$, is the following:
\begin{align*}
    \sigma, i &\models_D\ expression &\text{iff} \quad &1 \leq i \leq |\sigma|\ \text{and}\ evaluate(expression)(\sigma(i)) = true \\
    \sigma, i &\models_D\ \underline{\text{true}} \\
    \sigma, i &\not\models_D\ \underline{\text{false}} \\
    \sigma, i &\models_D \neg F &\text{iff} \quad &\sigma, i \not\models_D F \\
    \sigma, i &\models_D F_1 \land F_2 &\text{iff} \quad &\sigma, i \models_D F_1\ \text{and}\ \sigma, i \models_D F_2 \\
    \sigma, i &\models_D F_1 \lor F_2 &\text{iff} \quad &\sigma, i \models_D F_1\ \text{or}\ \sigma, i \models_D F_2 \\
    \sigma, i &\models_D F_1 \rightarrow F_2 &\text{iff} \quad &\sigma, i \models_D F_1 \text{ implies } \sigma, i \models_D F_2 \\
    \sigma, i &\models_D \bigcirc F &\text{iff} \quad &i \leq |\sigma|\ \text{and}\ \sigma, i+1 \models_D F \\
    \sigma, i &\models_D \Bigg\bigodot F &\text{iff} \quad &1 \leq i\ \text{and}\ \sigma, i-1 \models_D F \\
    \sigma, i &\models_D F_1 \cdot F_2 &\text{iff} \quad &\exists j \text{ such that } i \leq j \leq |\sigma|+1 \quad \text{and} \\
    & &&\sigma[1, j-1], i \models_D F_1 \quad \text{and} \quad \sigma[j,|\sigma|], 1 \models_D F_2 \\
    \sigma, i &\models_D N(F, P) &\text{iff} 
    &\begin{cases}
    \text{if } 1 \leq i \leq |\sigma| \text{ then:} \\
    \sigma, i \models_D F[ f \mapsto F, p \mapsto \text{evaluate}(P)(\sigma(i))] \\
    \text{where } (N(\text{Form } f, \text{T } p) = F) \in D \\
    \text{otherwise, if } i = 0 \text{ or } i = |\sigma|+1 \text{ then:} \\
    \text{rule } N \text{ is defined as max in } D
    \end{cases}
\end{align*}

>>>>

As we can see, rules are parameterized by an EAGLE formula (of type $\underline{\text{Form}}$), which means that we will be able to write EAGLE formulas such as $\text{Always}(\text{Eventually}(x>0))$. Also, the $\text{Always}$ operator is defined as maximal solution of the equation $X=F\land \bigcirc X$ and the $\text{Eventually}$ operator represents the minimal solution to the equation $X=F\lor \bigcirc X$.

@@

Given a trace $\sigma$ and a specification $D \, O$, satisfaction is defined as follows:
\[
\sigma \models D \, O \text{ iff } \forall (\underline{\text{mon}}\ N = F) \in O.\ \sigma ,1 \models_D \, F
\]
In other words, a trace satisfies a specification if it fulfills each monitored formula when monitoring from position 1, which is the initial state. An \textbf{important} note is that, given a trace $\sigma=\alpha_{1},\ldots ,\alpha_{n}$ the index $i$ of the trace can become $0$ (before the first state) and $n+1$, thus going beyond the limits; both of these situations result in rule applications evaluating to either true if maximal or false if minimal, without first taking the rules body into account. The definition of the satisfaction relation $\models_D \subseteq (Trace \times \mathbb{N}) \times \underline{\text{Form}}$, for a set of rule definitions $D$, is the following:
\begin{align*}
    \sigma, i &\models_D\ expression &\text{iff} \quad &1 \leq i \leq |\sigma|\ \text{and}\ evaluate(expression)(\sigma(i)) = true \\
    \sigma, i &\models_D\ \underline{\text{true}} \\
    \sigma, i &\not\models_D\ \underline{\text{false}} \\
    \sigma, i &\models_D \neg F &\text{iff} \quad &\sigma, i \not\models_D F \\
    \sigma, i &\models_D F_1 \land F_2 &\text{iff} \quad &\sigma, i \models_D F_1\ \text{and}\ \sigma, i \models_D F_2 \\
    \sigma, i &\models_D F_1 \lor F_2 &\text{iff} \quad &\sigma, i \models_D F_1\ \text{or}\ \sigma, i \models_D F_2 \\
    \sigma, i &\models_D F_1 \rightarrow F_2 &\text{iff} \quad &\sigma, i \models_D F_1 \text{ implies } \sigma, i \models_D F_2 \\
    \sigma, i &\models_D \bigcirc F &\text{iff} \quad &i \leq |\sigma|\ \text{and}\ \sigma, i+1 \models_D F \\
    \sigma, i &\models_D \Bigg\bigodot F &\text{iff} \quad &1 \leq i\ \text{and}\ \sigma, i-1 \models_D F \\
    \sigma, i &\models_D F_1 \cdot F_2 &\text{iff} \quad &\exists j \text{ s.t. } i \leq j \leq |\sigma|+1\ \text{and}\ \sigma^{[1, j-1]}, i \models_D F_1\ \text{and}\ \sigma^{[j,|\sigma|]}, 1 \models_D F_2 \\
    \sigma, i &\models_D N(\overline{F}, \overline{P}) &\text{iff} 
    &\begin{cases}
    \text{if } 1 \leq i \leq |\sigma| \text{ then:} \\
    \quad \sigma, i \models_D F[ \overline{f} \mapsto \overline{F}, \overline{p} \mapsto\ evaluate(\overline{P})(\sigma(i))] \\
    \quad \text{where } (N(\underline{\overline{\text{Form}}}\ \overline{f}, \text{T } p) = F) \in D \\
    \text{otherwise, if } i = 0 \text{ or } i = |\sigma|+1 \text{ then:} \\
    \quad \text{rule } N \text{ is defined as }\underline{\text{max}}\text{ in } D
    \end{cases}
\end{align*}
